Suicide Rate across the world

```{r knitr_setup, echo=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```
## Suicide Rate across the world

# Team member: 1 person
# Yang Lu

# Suicide rate across the world from 1985 to 2016

##1.Introduction

Suicide, it is a disaster that happen on people all over the world. Some people may have very bad memories about the suicide because some of the famous people died because of that. WIth the development of technology, we want to focus more on the mental health and reduce the problems of suicide in our society.

In order to investigate on the suicide and help reduce this kind of tragedy from happening, we need to know the statistics for that and know how many people in each country are suffering from that every year. Because one's suicide is related with their entire family, mitigating suicide rate can have a significant impact in our world.
 
Data science helped researchers to find out the correlation between suicide rate and its country to see whether the wealth of a country and how they can affect a country's suicide rate and what are other factors that may affect it.We are also curious on finding what are the countries with the lowest suicide rate and the highest suicide as well to see what are the characteristics of such countries and help us better understand the phenomenon.

Here we explore a dataset regarding the suicide rate from all over the world from 1985 to 2016, while also providing a walk through tutorial of the entire data science pipeline.


Dataset URL: https://www.kaggle.com/russellyates88/suicide-rates-overview-1985-to-2016/version/1

##Motivation
I am motivated to do this topic because citizens' health situation is the priority for every country in nowadays society. In addition to that, the development of technology should serve for the people in our world. As one of the biggest problem our society is facing that cause a huge amount of deaths, it is important for data science as this is useful for finding out the reasons behind the phenomenon and solve the problem in the future and this is the foundation data science stands for which is to help the society and solve the problems.

## 2.Getting Started with the Data

First, we are calling numerous libraries that are necessary for our code because these libraries 
consist of many useful R functions that will be used in the code.

## Required Libraries
```{r load_data, message=FALSE}
# Calling Libraries
library(tidyverse)
library(rvest)
library(dplyr)
library(ggplot2)
library(broom)
library(magrittr)
library(tidyr)
library(stringr)
```

## Importing the Data

After downloading a CSV file from Kaggle, use the read_csv() function with the file path as the 
parameter to read the data from a CSV file into a data frame in R. 
When loading a CSV file, you can parse it with column specifications using the cols() or 
cols_only() function. cols() includes all columns in the data, whereas cols_only() only include 
the columns specified.

For more information on the read_csv() function:   
  - https://readr.tidyverse.org/reference/read_delim.html 
for more information on the cols() and cols_only() function:
  - https://www.rdocumentation.org/packages/readr/versions/1.3.1/topics/cols 
Find datasets through links provided here: 
  - http://www.hcbravo.org/IntroDataSci/resources/ 
  - https://www.kaggle.com/datasets 
  
```{r}
csv_file <- "master.csv"
df <- read_csv(csv_file)
# Parsed with column specification
cols(
  country = col_character(),
  year = col_double(),
  sex = col_character(),
  age = col_character(),
  suicides_no = col_double(),
  population = col_double(),
  suicidesper100pop = col_integer(),
  country_year = col_character(),
  HDI = col_character(),
  gdp_for_year = col_double()
)
df
```

The set_colnames function allows for renaming column names.

```{r}
# Change column names
df <- df %>%
  set_colnames(c("Country", "year", "sex", "age range", "suicide_number", "total_population", "suicide_rate", "country_year","HDI","gdp_year","gdp_capita","generation"))
df
```

# Exploratory Data Analysis

Operations, such as select, filter, slice, arrange, group_by, and summarise, are used to help 
perform almost any analysis on data frames.

Using the various operations, we can determine which state has had the largest and smallest total
average honey production between 1998 and 2012, and on average, how many colonies it takes to 
create 1 pound of honey for each state.

Documentation of operations used:
  - group_by(): https://www.rdocumentation.org/packages/dplyr/versions/0.7.8/topics/group_by
  - summarise(): https://www.rdocumentation.org/packages/dplyr/versions/0.7.8/topics/summarise
  - filter(): https://www.rdocumentation.org/packages/dplyr/versions/0.7.8/topics/filter
  - mutate(): https://www.rdocumentation.org/packages/dplyr/versions/0.5.0/topics/mutate 


URL for the problem of Lithuania: https://en.wikipedia.org/wiki/Suicide_in_Lithuania

We can see economic problem and social transition has been a great factor which cause LIthuania to be so depressed. From this chart, we can clearly see that a country's economic situation is positively related to its suicide rate.
```{r}
# Lithuania has the highest suicide rate among all countries
df %>%
  group_by(Country) %>%
  summarise(avg_suiciderate = mean(suicide_rate)) %>%
  filter(avg_suiciderate == max(avg_suiciderate))
```

```{r}
# Dominica and Saint Kitts and Nevis has the lowest suicide rates
df %>%
  group_by(Country) %>%
 summarise(avg_suiciderate = mean(suicide_rate)) %>%
   filter(avg_suiciderate == min(avg_suiciderate))
```


```{r}
# On average, the amount of gpd shared by one person in the country and its suicide data
df %>%
  group_by(Country) %>%
  mutate(mean_gdp = gdp_year/total_population)
```

The plot below shows the suicide rate in each continent and it uses plots to help readers better understand the graph.

geom_point() is used to create the scatterplot and geom_smooth() is used to create the trend 
line. labs() and ggtitle() is used to customize the names of the axises and title of the plot.
In addition, theme(plot.title = element_text(hjust = 0.5)) helps center the title of the plot.

Here is a ggplot2 reference sheet: https://ggplot2.tidyverse.org/reference/ 

# Below are some countries from North America
```{r}
df %>%
  group_by(Country) %>%
 summarise(avg_suiciderate = mean(suicide_rate))      %>%
  filter(Country %in% c("United States", "Canada","Mexico" )) %>%
  ggplot(aes(x = Country, y = avg_suiciderate, color = Country)) + geom_col() + 
  geom_smooth(method = loess) + labs(x = "Country", y = "Suicide number") +
  ggtitle("Number of people commit suicide per 100k people") + 
  theme(plot.title = element_text(hjust = 1.0))
```

# Below are countries from Asia
```{r}
df %>%
  group_by(Country) %>%
 summarise(avg_suiciderate = mean(suicide_rate))      %>%
  filter(Country %in% c("Republic of Korea", "Japan","Thailand","Singapore" )) %>%
  ggplot(aes(x = Country, y = avg_suiciderate, color = Country)) + geom_col() + 
  geom_smooth(method = loess) + labs(x = "Country", y = "Suicide number") +
  ggtitle("Number of people commit suicide per 100k people") + 
  theme(plot.title = element_text(hjust = 1.0))
```

# Below are countries from Europe
```{r}
df %>%
  group_by(Country) %>%
 summarise(avg_suiciderate = mean(suicide_rate))      %>%
  filter(Country %in% c("Russian Federation", "France","Germany","United Kingdom","Italy" )) %>%
  ggplot(aes(x = Country, y = avg_suiciderate, color = Country)) + geom_col() + 
  geom_smooth(method = loess) + labs(x = "Country", y = "Suicide number") +
  ggtitle("Number of people commit suicide per 100k people") + 
  theme(plot.title = element_text(hjust = 1.0))
```


```{r}
df %>%
  group_by(Country) %>%
 summarise(avg_suiciderate = mean(suicide_rate))      %>%
  filter(Country %in% c("Australia", "New Zealand")) %>%
  ggplot(aes(x = Country, y = avg_suiciderate, color = Country)) + geom_col() + 
  geom_smooth(method = loess) + labs(x = "Country", y = "Suicide number") +
  ggtitle("Number of people commit suicide per 100k people") + 
  theme(plot.title = element_text(hjust = 1.0))
```

```{r}
df %>%
  group_by(Country) %>%
 summarise(avg_suiciderate = mean(suicide_rate))      %>%
  filter(Country %in% c("South Africa")) %>%
  ggplot(aes(x = Country, y = avg_suiciderate, color = Country)) + geom_col() + 
  geom_smooth(method = loess) + labs(x = "Country", y = "Suicide number") +
  ggtitle("Number of people commit suicide per 100k people") + 
  theme(plot.title = element_text(hjust = 1.0))
```


# Linear Regression

Linear regression is a very useful technique for data analysis.
It allows for constructing confidence intervals, utilizing hypothesis testing for relationships 
between variables, and providing continuous outcomes of interest.

From http://r-statistics.co/Linear-Regression.html , 
  "linear regression is used to predict the value of an outcome variable Y based on one or more 
  predictor variables X. The aim is to establish a linear relationship (a mathematical formula) 
  between the predictor variable(s) and the response variable, so that, we can use this formula 
  to estimate the value of the response Y, when only the predictors (Xs) values are known."

For more information on Linear Regression and Linear Models:
  - https://www.statisticssolutions.com/what-is-linear-regression/
  - http://www.stat.yale.edu/Courses/1997-98/101/linreg.htm 
  - https://data.princeton.edu/r/linearmodels 
      - This link helps explains how to fit a model, examine a fit, extract results, and much   
      more.

```{r}
# Fit a linear regression model for Suicide rate vs. Year
df_fit_suicide <- lm(suicide_rate~year, data = df)
df_fit_suicide
df_fit_suicide_stats <- df_fit_suicide %>%
  tidy()
df_fit_suicide_stats
``` 

```{r}
cat("On average, the number of people commit suicide amoung 100k people decreased by", df_fit_suicide_stats$estimate[2], 
    "pounds per year from 1985 to 2016.")
``` 
  

If there was a null hypothesis of no relationship between suicide rate and year, I 
would reject that null hypothesis because there is a relationship between the two variables: Over
time, the total suicide rate decreases by around -0.874 per year.

Now, we augment the df_fit_suicide, in which columns, such as predictions, residuals, etc. are 
added.
Residuals is the difference between the observed value of the dependent variable and the 
predicted value. Fitted values are also known as predicted values.
One way to check if your linear regression model is appropriate is to plot a graph of Residuals 
vs. Fitted Values. This graph will check for the linearity assumption. If the regression model is
appropriate, the mean of residuals will be approximately 0.

The augment() function Documentation: 
  - https://www.rdocumentation.org/packages/broom/versions/0.4.3/topics/augment 
Residuals:
  - http://www.r-tutor.com/elementary-statistics/simple-linear-regression/residual-plot 
  
```{r}
aug_df_suicide <- df_fit_suicide %>%
  augment()
aug_df_suicide
``` 


```{r}
# A violin plot of model Residuals vs. Year
aug_df_suicide %>%
  ggplot(aes(x = factor(year), y = .resid)) + geom_violin() + 
  labs(x = "Year", y = "Residuals") + ggtitle("Residuals vs. Year") + 
  theme(plot.title = element_text(hjust = 0.5))
``` 

## Conclusion
The suicide rate is positively related with a country's financial situation. We can see Lithuania is one of the poorest country in the world and it has a highest suicide rate in the world. However, once the economy level of a country reaches a certain level, we can see that the suicide problem is not only related to the economy situation of the country.

In my experienment I find out that Australia is among the countries with the lowest suicide rate, I conclude that it is not only related with financial situation but also with the living standard of the country.
URL: https://degreesplusaustralia.com.au/why-australia/living-standards/
WE can see from this article that Australia's people's costs of livving is low and they have a very high standard of living, they do not need to work for many hours and they can afford their family easily. 

This means they do not have many things to be bothered and that is the direct reason why Australians have the lowest suicide rate in the world.